---
title: "Exploratory analysis report for PUBG"
subtitle: "Kaggle project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,cache = TRUE,warning = F)

# mainPath<-file.path(Sys.getenv("R_USER"), "PUBG_Rproject", "scripts")
mainPath<-file.path(Sys.getenv("HOME"),"kaggle","PUBG_Rproject", "scripts")
setwd(mainPath)#"~/kaggle/PUBG_R/PUBG_R/"


dataDir<-file.path(getwd(), "data");

# install external libraries
library(dplyr);library(ggplot2);library(reshape2);#library(ggpubr)#library(gridExtra)
# library(data.table)
```

### Synopsis

PUBG is a successfull multiplayer online game set on an island where one hundred people play a battle royal game. 
Each player struggles in fights, hidings, escapes, pursuits, ambushes, stalkings and whatever mean necessary in order to survive. 
Only the last player standing wins. Fight or escape? Hide or attack? Join forces or play alone?
In order to decide the best strategy, a set of player stats have been collected and analyzed. 
The final aim is to apply a predictive model that target the player ranking.

## Data Set

PUBG is organized in matches. In each match, up to 100 people partecipate as singles or teams. At the end of the match, 
players (singles or groups) are ranked based on how many other groups are still alive when they are eliminated.
During the match, players can find objects such as weapons, vehicles, and medical kits they can use to kill and injure 
other players (including team members), to drive away from the dangerous areas or go right in to the middle of it.

```{r load data}
dataTrainSam<-read.csv(file.path(dataDir,"train_V2.csv"),stringsAsFactors =F,nrows=10)
dataTrain<-read.csv(file.path(dataDir,"train_V2.csv"),stringsAsFactors =F, 
                    col.names = names(dataTrainSam),colClasses = sapply(dataTrainSam,class))
rm(dataTrainSam)
dimData<-dim(dataTrain)
sizeData<-format(object.size(dataTrain),units = "Gb")

cutpoints<-c(0,0.2,0.9,0.999999999,1.01);
lab=c("tail","heart","head","top") 
scores<-cut(dataTrain$winPlacePerc,breaks=cutpoints,labels =lab,right = F ) #[a,b)

rm(dataTrain)
```

The data set, available [here](https://www.kaggle.com/c/pubg-finish-placement-prediction/data), 
consists of `r dimData[2]` statistics collected for `r dimData[1]` players weighting
`r sizeData`. The statistics are:

1. **Id**: player’s Id
2. **assists**: number of enemy this player damaged that were killed by teammates
3. **boosts**: number of boost items used
4. **heals**: number of healing items used
5. **revives**: number of times this player revived teammates
6. **damageDealt**: total damage dealt. Note: Self inflicted damage is subtracted
7. **DBNOs**: number of enemy knocked
8. **killPlace**: ranking in match of number of enemy killed
9. **killPoints**: kills-based external ranking of player ["0" should be treated as a “None” 
for *rankPoints* equal to -1]
10. **killStreaks**: max number of enemy killed in a short amount of time
11. **kills**: number of enemy killed
12. **headshotKills**: number of enemy killed with headshots
13. **roadKills**: number of kills while in a vehicle
14. **teamKills**: number of times this player killed a teammate
15. **longestKill**: longest distance between player and player killed at time of death
16. **rideDistance**: total distance traveled in vehicles measured in meters
17. **swimDistance**: total distance traveled by swimming measured in meters
18. **vehicleDestroys**: number of vehicles destroyed
19. **walkDistance**: total distance traveled on foot measured in meters
20. **weaponsAcquired**: number of weapons picked up
21. **winPoints**: win-based external ranking of playe ["0" should be treated as a “None” for
*rankPoints* equal to -1]
22. **winPlacePerc**: percentile winning placement, where 1 corresponds to 1st place. 
It is calculated off of *maxPlace* [TARGET]
23. **rankPoints**: Elo-like ranking of player, inconsistent and is being deprecated in the API’s next version
24. **groupId**: ID group within a match. In different matches the same group of players will have a different IDs
25. **matchDuration**: duration of match in seconds
26. **matchId**: ID to identify match
27. **matchType**: game mode such as: “solo”, “duo”, “squad”, “solo-fpp”, “duo-fpp”, “squad-fpp”, and other custom modes
28. **numGroups**: number of groups we have data for in the match
29. **maxPlace**: worst placement we have data for in the match

Since the big number of statistics and observations, summary plots for match and single players features are shown.

## Summary statistics

### Match statistics


``` {r data_match}
# dataMatch<-dataTrain %>% mutate(score=scores=="top") %>% add_count(matchId)  %>% 
#    mutate(nPlayers=n) %>% select(-n) %>%  group_by(matchId) %>% summarise(type=unique(matchType,matchId),duration=mean(matchDuration),nGroups=mean(numGroups), 
#                                   maxPlace=mean(maxPlace),nPlayers=unique(nPlayers,matchId),nWinners=sum(score))

dataMatch<-read.csv(file=file.path("working_data","dataMatch.csv"))
summMatch<-dataMatch %>% summarise(AVduration=mean(duration),SDduration=sd(duration),
                                   AVnGroups=mean(nGroups), SDnGroups=sd(nGroups),
                                  AVmaxPlace=mean(maxPlace),SDmaxPlace=sd(maxPlace),
                                  AVnPlayers=mean(nPlayers), SDnPlayers=sd(nPlayers),
                                  AVnWinners=mean(nWinners,na.rm = T),SDnWinners=sd(nWinners,na.rm = T))

```


Data contains statistics of `r nrow(dataMatch)` matches. For each match it is reported: the type 
[`r length(unique(dataMatch$type))` types], the duration, the number of partecipating players, the number of partecipating groups, 
the worst placement ranked in the match, and the number of winners in the match.  
Although the match number is not even amongs types, as shown in the table below, notice that there is a high variability in 
the statistic distributions for each match type [see Figure 1]. For this reason the match type is disregarded in the next analysis.

```{r n_types,tidy=T}
 NMtype<-table(dataMatch$type)
 print(NMtype)
```

```{r data_match_plot1}
gm1<-ggplot(data=dataMatch,aes(y=duration,fill=type))+geom_boxplot()+ylab("duration[s]")+
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(),axis.ticks.x=element_blank())
gm2<-ggplot(data=dataMatch,aes(y=nPlayers,fill=type))+geom_boxplot()+ylab("N players")+
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(),axis.ticks.x=element_blank())
gm3<-ggplot(data=dataMatch,aes(y=nGroups,fill=type))+geom_boxplot()+ylab("N groups")+
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(),axis.ticks.x=element_blank())
gm4<-ggplot(data=dataMatch,aes(y=maxPlace,fill=type))+geom_boxplot()+ylab("worst placement")+
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(),axis.ticks.x=element_blank())
gm5<-ggplot(data=dataMatch,aes(y=nWinners,fill=type))+geom_boxplot()+ylab("N winners")+
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(),axis.ticks.x=element_blank())

fig1<-ggarrange(gm1,gm2,gm3,gm4,gm5,nrow=5,common.legend=TRUE,legend="right")
annotate_figure(fig1, top = "Match statistics", fig.lab = "Figure 1", fig.lab.pos = "bottom")
rm(fig1)
```

On average, a match lasts `r format(summMatch$AVduration,digits=4)` +/- `r format(summMatch$SDduration,digits=4)` s
[~`r format(summMatch$AVduration/60,digits=2)`  +/- `r format(summMatch$SDduration/60,digits=2)` min]
and has `r format(summMatch$AVnPlayers,digits=2)` +/- `r format(summMatch$SDnPlayers,digits=2)` players divided in 
`r format(summMatch$AVnGroups,digits=4)` +/- `r format(summMatch$SDnGroups,digits=4)` groups. 
At the end of the match, on average, the worst placement is `r format(summMatch$AVmaxPlace,digits=4)` +/- 
`r format(summMatch$SDmaxPlace,digits=4)` and the number of winners is `r format(summMatch$AVnWinners,digits=1)` +/-
`r format(summMatch$SDnWinners,digits=1)`. 

### Personal statistics

For a better visualization, the over 4 milions players have been divided into 4 groups depending on their placement at the end 
on the match, i.e. the **winPlacePerc** spanning between 0 and 1. Groups represent winning, good, medium, and poor performances 
rather than an uniform partition of the players. Winning players score 1 and are labeled as "top"; players 
scoring between 0.9 and 0.99 are labeld as "head", while scores between 0.2 and 0.9 (excluded) are labeld as "heart" and 
scores between 0 and 0.2 (excluded) are labeld as "tail" (see table below).


```{r score_def1}

dfScore<-data.frame(score=c("[0-0.2)","[0.2-0.9)","[0.9-0.99]","1"),number=as.integer(table(scores)),
                    percentage=format(table(scores)/length(scores),digits=1))
print(dfScore)
```

For each class, the value distribution of the player statistics are shown as histograms in Figure 2. 

```{r players_stats}
# dataPlayerS<-dataTrain %>% 
#   mutate(DBNOs=DBNOs,damageDealt=damageDealt,weaponsAcquired=weaponsAcquired,vehicleDestroys=vehicleDestroys,
#          killPoints=killPoints,killPlace=killPlace,longestKill=longestKill,
#          totKills=headshotKills+killStreaks+roadKills+kills+teamKills,
#          totDistance=swimDistance+walkDistance+rideDistance,
#          heals=heals,revives=revives,assists=assists,boosts=boosts,
#          winPoints=winPoints,score=scores) %>% 
#   select(Id,groupId,DBNOs,damageDealt,weaponsAcquired,killPoints,killPlace,longestKill,
#          totKills,totDistance,
#          vehicleDestroys,
#          heals,revives,assists,boosts,winPoints,score)
dataPlayerS<-read.csv(file=file.path("working_data","dataPlayerS.csv"))
```


```{r players_stats_hist}
playerVariables<-c("DBNOs","damageDealt","weaponsAcquired","killPoints","killPlace","longestKill",
                   "totKills","totDistance","vehicleDestroys", "heals","revives","assists","boosts","winPoints")

g_glob<-list()

# for(i in c(1,2,3,4,5)){#seq_along(playerVariables)
#   rangeVar<-range(dataPlayerS[,playerVariables[i]])
#   g_glob[[i]]<-ggplot(data = dataPlayerS,aes(x=dataPlayerS[,playerVariables[i]],fill=score,stat(ncount))) +
#             geom_histogram(binwidth=max(rangeVar)/10,position = "dodge") +coord_cartesian(xlim = rangeVar)+
#              xlab(playerVariables[i])+ ylab("% of max")
# }

# fig2<-ggarrange(g_glob[[1]],g_glob[[2]],g_glob[[3]],g_glob[[4]],g_glob[[5]],
#              g_glob[[6]],g_glob[[7]],g_glob[[8]],g_glob[[9]],g_glob[[10]],
#              g_glob[[11]],g_glob[[12]],g_glob[[13]],g_glob[[14]],
# nrow=4,common.legend=TRUE,legend="right")
# annotate_figure(fig2, top = "Player statistics", fig.lab = "Figure 2", fig.lab.pos = "bottom")
```

